\subsection{User Story \#NUMBER: Workspace for Term and Glossary Management}

\textbf{ID:} US002 (Marito Project) \\
\textbf{Title:} Personalized Workspace for Managing Terms and Glossaries \\
\textbf{As a:} Marito application user (e.g., linguist, researcher, contributer), \\
\textbf{I want:} a workspace where I can view my saved terms, track the progress of submitted terms, view glossaries or languages I've saved, and organize them into named groups (e.g., ``Thesis Research''), \\
\textbf{So that:} I can efficiently manage and organize my linguistic research and contributions in one place.

\vspace{1em}
\textbf{Acceptance Criteria:}
\begin{enumerate}
    \item \textbf{View Saved Terms:}
    \begin{itemize}
        \item \textbf{Given} I have saved terms,
        \item \textbf{Then} I can view them in a dedicated workspace section, with options to search and filter.
    \end{itemize}

    \item \textbf{Track Submitted Term Progress:}
    \begin{itemize}
        \item \textbf{Given} I have submitted terms for review,
        \item \textbf{Then} I can see their current status (e.g., ``Under Review'', ``Approved'', ``Rejected'') in my workspace.
    \end{itemize}

    \item \textbf{View Saved Glossaries or Languages:}
    \begin{itemize}
        \item \textbf{Given} I have saved glossaries or language sets,
        \item \textbf{Then} I can view them in my workspace for quick access.
    \end{itemize}

    \item \textbf{Group Terms and Glossaries:}
    \begin{itemize}
        \item \textbf{Given} I want to organize my saved items,
        \item \textbf{Then} I can create named groups (e.g., ``Thesis Research'') and assign terms or glossaries to these groups.
    \end{itemize}

    \item \textbf{Manage Groups:}
    \begin{itemize}
        \item \textbf{Then} I can rename or delete groups,
        \item \textbf{And} I can move items between groups.
    \end{itemize}

    \item \textbf{Offline Accessibility:}
    \begin{itemize}
        \item \textbf{Given} I am offline,
        \item \textbf{Then} I can still access my saved workspace data including saved terms and glossaries.
    \end{itemize}

    \item \textbf{Data Syncing:}
    \begin{itemize}
        \item \textbf{Given} I go online,
        \item \textbf{Then} any changes to saved data or submission statuses are synchronized automatically with the central repository.
    \end{itemize}

    \item \textbf{Mobile-Responsive UI:}
    \begin{itemize}
        \item \textbf{Then} the workspace UI is optimized for mobile, tablet, and desktop views.
    \end{itemize}
\end{enumerate}

\vspace{1em}
\textbf{Notes/Assumptions:}
\begin{itemize}
    \item Groups act like folders or collections and are only visible to the user who created them.
    \item Saved term statuses are maintained on the backend and synced upon login or reconnection.
    \item Offline access will use local caching or IndexedDB mechanisms.
    \item Glossary or term contributions are assumed to be submitted while online.
\end{itemize}
